Developing a preprint culture in biology

### 2012-10-05 ms.tex
TP funding

%It is unclear to me whether we should keep this full paragraph or try to distill
%it to a sentence or two at the end of the previous paragraph.
Furthermore, the formal peer-review process as a whole is critically over-loaded. As
the number of active scientists increases and the pressure to publish
increases, it has become increasingly difficult for journals to find
reviewers \cite{hoc09}.  At the same time, rejection rates are high in most journals
\cite{aar08,roh09}, and when not invited to submit a revision, authors must start the
process over again at another journal. As a result, initiatives to reduce time from submission to
publication have emerged across the scientific community. Rohr et al.
\cite{roh09} called for the recycling and reuse of peer-reviews:
by attaching previous reviews and detailed replies to a new submission, both
the editor and the referees can gauge the work done on the manuscript, and
perhaps evaluate it with less prejudice. In a similar way, the \emph{Peerage
of Science} initiative allows authors to seek anonymous pre-review by their
peers. Some journals now accept to publish papers which received good
evaluations, with \emph{Animal Biology} having recently accepted first a paper
reviewed entirely with the \emph{Peerage of Science} \cite{abb12}, effectively
outsourcing the review process. A widespread use of preprint servers can
achieve the same goal of reducing the time spent in review. By allowing
open comments and criticisms prior to submission, the authors will receive valuable
feedback and can improve the version which will be submitted. With a rich
enough community of scientists depositing preprints, and commenting on them,
the process of an open pre-review can become widespread and will overall
increase the quality of first submissions.

@@

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice. This is 
based on a certain interpretation
of the Ingelfinger rule: scientists should not publish the same manuscript twice
\cite{alt96}. However, preprints are not peer-reviewed papers and posting to a
public server does not constitute publication in a peer-reviewed journal.
A preprint is simply a document that allows ideas to spread and be discussed, 
but it is not yet formally validated by the peer-review system. This is why the
majority of publishers do not see arXiv and similar services as a violation of the
Ingelfinger rule. Almost all of the major publishers in biology are preprint-friendly,
including: Nature Publishing Group, PLOS, BMC, PNAS, Science (mostly) \ref{table:policies},
and all the journals from Elsevier %something is off in the formatting here
and Springer.  The Ecological Society of America and the Genetics Society of
America also recently changed their policies to allow public preprints.  Few
scientific publications will not consider a manuscript submitted to a public
preprint server. In fact, \emph{Nature} responded to the rumour that they refused
manuscripts submitted to arXiv by saying that ``\emph{Nature} never wishes to
stand in the way of communication between researchers. We seek rather to add
value for authors and the community at large in our peer review, selection and
editing'' \cite{nat05}. Still, a few journals adopt a ``by default'' hostile attitude
towards preprints, mostly due to the lack of clear policy of the publishers, or
perhaps because a preprint culture has not developed in biology and the practice
is still considered unusual. As an example, Wiley-Blackwell, which publishes
some of the leading journals in biology, has no official policy on the matter.

\ref{table:policies}.

@@

% Dominique:
% I added the following three paragraphs to the conclusion, offering also a
% discussion on the on going explosion of different publication models. Take
% what you want and remove what you don't need.  I also had the feeling we might
% want to include a single paragraph, very short, calling that evaluations
% (academia, grant agencies etc...) looking at publication records or report of
% scientific activities start to view preprints as a significant and sound
% scientific contribution. One of the most important benefit is again the speed
% of the publication process. Most grants are for three years or less. It makes
% it hard to have a published paper within the limits of the grant. Preprints
% could however speed the process and show that contributions were made during
% the time of the grant.

@@

% Aren't we bridging more than cultural and geographic? I mean, we are
% essentially cross pollinating ideas. e.g. It took a lot time for MCMC to
% become popular in ecology even though it was widely used in other disciplines
% a decade or two earlier.
Preprint servers also facilitate communication between disciplines, bridging
cultural as well as geographic divides. Examples include a recent series of
papers on the theory of natural selection that was posted to arXiv
simultaneously with its publication in the \emph{Journal of Evolutionary
Biology} \cite{JEB:JEB2431,JEB:JEB2498,JEB:JEB2378,JEB:JEB2373}. Other
submissions in this category include evolutionary and ecological theory by
authors trained in physics and computer science.  Since authors in these fields
regularly check arXiv, submitting preprints may be the most effective way
biologists can help others avoid ``repeated work'' \cite{de2011contribution} and
form a synthetic community of evolutionary theorists from disparate backgrounds.
The advantages are clear and the costs are low.

@@

PDP is supported by an Alexander Graham Bell scholarship from the National
Sciences and Engineering Council of Canada. EPW is supported by a CAREER Award
from the National Science Foundation (DEB-0953694). JJA is supported by NSF
DEB-0614166 and NSF DEB-0919018. KR is supported by NSF DEB-1021553. DG is
funded by a Discovery Grand from the National Sciences and Engineering Council
of Canada and by the Canada Research Chair program.

>>>>

Furthermore, the formal peer-review process as a whole is critically
over-loaded. As the number of active scientists increases and the pressure to
publish increases, it has become increasingly difficult for journals to find
reviewers \cite{hoc09}.  At the same time, rejection rates are high in most
journals \cite{aar08,roh09}, and when not invited to submit a revision, authors
must start the process over again at another journal. As a result, initiatives
to reduce time from submission to publication have emerged across the scientific
community. Rohr et al.  \cite{roh09} called for the recycling and reuse of
peer-reviews: by attaching previous reviews and detailed replies to a new
submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. In a similar way, the
\emph{Peerage of Science} initiative allows authors to seek anonymous pre-review
by their peers. Some journals now accept to publish papers which received good
evaluations, with \emph{Animal Biology} having recently accepted first a paper
reviewed entirely with the \emph{Peerage of Science} \cite{abb12}, effectively
outsourcing the review process. A widespread use of preprint servers can achieve
the same goal of reducing the time spent in review. By allowing open comments
and criticisms prior to submission, the authors will receive valuable feedback
and can improve the version which will be submitted. With a rich enough
community of scientists depositing preprints, and commenting on them, the
process of an open pre-review can become widespread and will overall increase
the quality of first submissions.

@@

Another major concern is that if a researcher posts a public preprint that they
will not be allowed to submit their paper to their journal of choice. This is
based on a certain interpretation of the Ingelfinger rule: scientists should not
publish the same manuscript twice \cite{alt96}. However, preprints are not
peer-reviewed papers and posting to a public server does not constitute
publication in a peer-reviewed journal.  A preprint is simply a document that
allows ideas to spread and be discussed, but it is not yet formally validated by
the peer-review system. This is why the majority of publishers do not see arXiv
and similar services as a violation of the Ingelfinger rule. Almost all of the
major publishers in biology are preprint-friendly, including: Nature Publishing
Group, PLOS, BMC, PNAS, Science (mostly) \ref{table:policies}, and all the
journals from Elsevier and Springer.  The Ecological Society of America and the
Genetics Society of America also recently changed their policies to allow public
preprints.  Few scientific publications will not consider a manuscript submitted
to a public preprint server. In fact, \emph{Nature} responded to the rumour that
they refused manuscripts submitted to arXiv by saying that ``\emph{Nature} never
wishes to stand in the way of communication between researchers. We seek rather
to add value for authors and the community at large in our peer review,
selection and editing'' \cite{nat05}. Still, a few journals adopt a ``by
default'' hostile attitude towards preprints, mostly due to the lack of clear
policy of the publishers, or perhaps because a preprint culture has not
developed in biology and the practice is still considered unusual. As an
example, Wiley-Blackwell, which publishes some of the leading journals in
biology, has no official policy on the matter.

@@

@@

@@

PDP is supported by an Alexander Graham Bell scholarship from the National
Sciences and Engineering Council of Canada. EPW is supported by a CAREER Award
from the National Science Foundation (DEB-0953694). JJA is supported by NSF
DEB-0614166 and NSF DEB-0919018.  TP is supported by a FQRNT-MELS post-doctoral
scholarship and 25 cents found by a coffee machine.  KR is supported by NSF
DEB-1021553. DG is funded by a Discovery Grand from the National Sciences and
Engineering Council of Canada and by the Canada Research Chair program.
